The importance of measurement uncertainty in ecological management

### 2018-01-31 manuscripts/manuscript.Rmd

To better understand the differences in performance of these strategies and resolve the paradox of uncertainty, we must take a closer look at how the specific action recommended by each policy compares given the same observation.  Figure \ref{policy} shows the action taken by each strategy in response to a measurement of the stock size (biomass estimate).  These plots show, for any possible observations, which strategy will attempt to harvest most and which will attempt to harvest least.  The first panel shows the policy in terms of expected escapement, $S = Y - H$ for observed stock $Y$ and harvest $H$, consistent with most of the relevant control and bioeconomics literature [e.g. @Reed1979; @Clark1986; @Sethi2005].  For comparison, the second panel shows the same policies but in terms of harvest $H$.  These plots clearly illustrate the contrast between the constant escapement (CE) strategy and the precautionary Total Allowable Catch (TAC) strategy: CE sets harvest strictly to zero for stocks estimated at biomass below $B_{MSY}$, while TAC permits a modest harvest of even very small stocks.  Escapement CE strategy corresponds to keeping escapement constant.  This is important, as theoretical work ever since the proofs of @Reed1979 often attempts to simplify the calculations by assuming the optimal policy must be a constant-escapement type strategy [e.g. @Ludwig1981; @Roughgarden1996; @Engen1997;  @Moxnes2003].  Note that both CE and TAC policies are independent of the levels of uncertainty $\sigma_m$ and $\sigma_g$.  TAC implicitly accounts for some uncertainty through its 80% MSY definition -- comparable plots with MSY can be found in the Supplemental Material.  In principle, the CE policy could depend on $\sigma_g$, but as @Reed1979 proved, the constant escapement level with stochasticity is the same as in the deterministic case, S = D, unless the noise level is quite high (comparison plots in Supplemental material.)  In contrast to this, our POMDP plots depend very much on both the choice of $\sigma_m$ and $\sigma_g$.  If $\sigma_m = 0$, they reduce exactly to the CE solution.  Here we show the corresponding policies for POMDP solutions focusing on $\sigma_g = 0.15$ as above, with a modest measurement uncertainty $\sigma_m = 0.1$.  Alternate combinations of measurement uncertainty can be found in the appendix but do not change the general pattern.  In addition to an explicit dependence on the measurement error, our POMDP solutions depend on another piece of information: any prior observations of the stock size.

Different attention to prior events fundamentally drives the differences between the POMDP strategy and the other strategies and can resolve the uncertainty paradox.  While the catch quota under both TAC and CE strategies can be completely determined given the most recent observation of the stock size by using the policy curves shown in Figure \ref{policy}, this is not the case for the POMDP approach. This fundamental difference is key to understanding the difference in performance and resolving the paradox of uncertainty.  The POMDP policy cannot be specified by the most recent observation alone. Instead, the POMDP policy depends on all prior observations, not just the most recent.  The reason for this complexity comes from the Markov property.  Observations of the state in the perfectly observed system satisfy the Markov property: once we have measured the current biomass exactly, we cannot get any better estimate of the current stock size by studying older measurements. When measurements are uncertain this is no longer the case: intuitively, by comparing the most recent measurement to previous observations we may be able to infer when any given measurement is unusually high or unusually low.  POMDP formalizes this intuition by capturing the information from all previous observations into a *prior belief*.  This prior belief is updated after every subsequent action and observation in accordance with Bayes Law.  The mechanics of this process are well documented in the extensive literature on POMDPs [e.g. @Smallwood1973; @Sondik1978; @Kurniawati2008; @Williams2011], but for our purposes it is sufficient to observe how this evolving prior belief serves to continually adjust the POMDP policy.  Figure S8 illustrates how the prior belief evolves in response to subsequent observations and actions over the course of an individual simulation.

@@

```{r historical, message=FALSE, warning=FALSE, fig.width=7, fig.height = 5, fig.cap="Comparisons of harvest level that would be recommended by policies considered here relative to historical harvest levels in two commercial fish stocks.  POMDP solution assumes a measurement error of 10%. \\label{historical}"}
ram_ex %>%
  filter(model %in% c("biomass", "catch", "POMDP", "CE", "TAC")) %>%
  ggplot(aes(year, stock, col=model)) +
  geom_line(lwd=1) +
  facet_wrap(~commonname, scales = "free", ncol=1)  
```

@@

These historical examples provide another useful lens to compare how the different strategies respond in the face of fluctuations in real data, rather than model simulations.  In both stocks, historical catch almost always exceeds that recommended by any of our strategies, falling closer to the MSY value estimated (see Appendix B).   In Argentine Hake, the persistent declines result in both POMDP and CE strategies quickly closing down the fishery in an attempt to let the stock recover to a more productive target biomass, while TAC persists with merely a reduced harvest.  A small recovery of biomass in 1995 is met by an immediate uptick in harvest under CE, while the POMDP response is more conservative.  The example of European Plaice illustrates more volatility in the stock size, revealing further differences in the strategies.  Despite this volatility, the TAC level remains relatively level across all five decades, rising or dipping only slightly with changes in stock size.  In contrast, the CE harvest tracks this volatility almost in perfect mirror.  Here, the POMDP solution falls in between these extremes, almost mirroring which-ever of the two policies is more conservative: the POMDP solution matches the dips in harvest taken by CE to allow the stock to recover most quickly, but does not track the doubling of harvests recommended by CE in the mid 1980, increasingly only to the more modest harvest recommended by the TAC policy.   Once again, we see that POMDP solution provides a more consistently precautionary policy than either the over-simplified optimal control solution of CE or the more rule-of-thumb approach represented by TAC.  Though we have seen that the POMDP solution need not always be more conservative (lower harvest) than these strategies, it makes use of the all prior observations to tune the level of caution appropriately.

>>>>

To better understand the differences in performance of these strategies and resolve the paradox of uncertainty, we must take a closer look at how the specific action recommended by each policy compares given the same observation.  Figure \ref{policy} shows the action taken by each strategy in response to a measurement of the stock size (biomass estimate).  These plots show, for any possible observations, which strategy will attempt to harvest most and which will attempt to harvest least. For comparison purposes, we plot policies both in terms of expected escapement, $S = Y - H$ as is typical in the optimal control literature, [e.g. @Reed1979; @Clark1986; @Sethi2005], and also directly in terms of the harvest quota $H$. These plots clearly illustrate the contrast between the constant escapement (CE) strategy and the precautionary Total Allowable Catch (TAC) strategy: CE sets harvest strictly to zero for stocks estimated at biomass below $B_{MSY}$, while TAC permits a modest harvest of even very small stocks. (Comparable plots with MSY can be found in the Supplemental Material).  In principle, the CE policy could depend on $\sigma_g$, but as @Reed1979 proved, the constant escapement level with stochasticity is the same as in the deterministic case, S = D, unless the noise level is quite high (comparison plots in Supplemental material.)  In contrast to this, our POMDP plots depend very much on both the choice of $\sigma_m$ and $\sigma_g$.  If $\sigma_m = 0$, they reduce exactly to the CE solution.  Here we show the corresponding policies for POMDP solutions focusing on $\sigma_g = 0.15$ as above, with a modest measurement uncertainty $\sigma_m = 0.1$.  Alternate combinations of measurement uncertainty can be found in the appendix but do not change the general pattern.  In addition to an explicit dependence on the measurement error, our POMDP solutions depend on another piece of information: any prior observations of the stock size.

Differences in attention to prior beliefs,determined by prior observations, drive the differences between the POMDP strategy and the other strategies and can resolve the uncertainty paradox.  While the catch quota under both TAC and CE strategies can be completely determined given the most recent observation of the stock size by using the policy curves shown in Figure \ref{policy}, this is not the case for the POMDP approach. This fundamental difference is key to understanding the difference in performance and resolving the paradox of uncertainty.  The POMDP policy cannot be specified by the most recent observation alone. Instead, the POMDP policy depends on all prior observations, not just the most recent.  The reason for this complexity comes from the Markov property.  Observations of the state in the perfectly observed system satisfy the Markov property: once we have measured the current biomass exactly, we cannot get any better estimate of the current stock size by studying older measurements. When measurements are uncertain this is no longer the case: intuitively, by comparing the most recent measurement to previous observations we may be able to infer when any given measurement is unusually high or unusually low.  POMDP formalizes this intuition by capturing the information from all previous observations into a *prior belief*.  This prior belief is updated after every subsequent action and observation in accordance with Bayes Law.  The mechanics of this process are well documented in the extensive literature on POMDPs [e.g. @Smallwood1973; @Sondik1978; @Kurniawati2008; @Williams2011], but for our purposes it is sufficient to observe how this evolving prior belief serves to continually adjust the POMDP policy.  Figure S8 illustrates how the prior belief evolves in response to subsequent observations and actions over the course of an individual simulation.

@@

```{r historical, message=FALSE, warning=FALSE, fig.width=7, fig.height = 5, fig.cap="Comparisons of harvest level that would be recommended by policies considered here relative to historical harvest levels in two commercial fish stocks.  POMDP solution assumes a measurement error of 10%. \\label{historical}"}
ram_ex %>%
  filter(model %in% c("biomass", "catch", "POMDP", "CE", "TAC")) %>%
  ggplot(aes(year, stock, col=model)) +
  geom_line(lwd=1) +
  facet_wrap(~commonname, scales = "free", ncol=1)  +
  ylab("stock biomass (metric tonnes)")
```

@@

These historical examples provide another useful lens to compare how the different strategies respond in the face of fluctuations in real data, rather than model simulations.  In both stocks, historical catch almost always exceeds that recommended by any of our strategies, falling closer to the MSY value estimated (see Appendix B).   In Argentine Hake, the persistent declines result in both POMDP and CE strategies quickly closing down the fishery in an attempt to let the stock recover to a more productive target biomass, while TAC persists with merely a reduced harvest.  A small recovery of biomass in 1995 is met by an immediate uptick in harvest under CE, while the POMDP response is more conservative.  The example of European Plaice illustrates more volatility in the stock size, revealing further differences in the strategies.  Despite this volatility, the TAC level remains relatively level across all five decades, rising or dipping only slightly with changes in stock size.  In contrast, the CE harvest tracks this volatility almost exactly.  Here, the POMDP solution falls in between these extremes, almost mirroring which-ever of the two policies is more conservative: the POMDP solution matches the dips in harvest taken by CE to allow the stock to recover most quickly, but does not track the doubling of harvests recommended by CE in the mid 1980, increasingly only to the more modest harvest recommended by the TAC policy.   Once again, we see that POMDP solution provides a more consistently precautionary policy than either the over-simplified optimal control solution of CE or the more rule-of-thumb approach represented by TAC.  Though we have seen that the POMDP solution need not always be more conservative (lower harvest) than these strategies, it makes use of the all prior observations to tune the level of caution appropriately.
